{\bf Answer:}

{\bf Part a:} \citep{kaplan2020scaling} is a study on neural scaling laws that investigates the effect of variables such as the model architecture, the model size (the network's width and depth for example), the computing power used to train, and the amount of training dataset on the overall performance. While related work exists that explore such trends in models such as random forests~\citep{biau2012analysis} and image models~\citep{tan2019efficientnet}, this study focuses on language models. The authors in~\citep{kaplan2020scaling} investigate and report power-law relationships between the performance (in terms of the loss on the test set) of the classical Transformer model~\citep{vaswani2017attention} on the WebText2 dataset~\citep{radford2019language} and three main variables of interest; the number of model parameters $N$, the dataset size $D$, and the amount of compute required $C$. Across the variety of the experiments conducted, the following trends are reported in the paper:
\begin{enumerate}
\item {\bf Power laws:} When not bottlenecked by the other two, the performance has a power-law relationship with each of the three factors $N, D, C$.
\item {\bf Overfitting:} If suboptimal values of $N$ or $D$ are used, the performance incurs a penalty as the scaling variable is increased.
\item {\bf Model size:} Large models tend to perform better overall; they need fewer optimization steps and a smaller dataset to achieve the same accuracy as smaller models. In fact, according to the authors the optimal performance is reached by training very large models and stopping significantly short of the convergence criteria.
\item {\bf Transfer learning:} Interestingly, the model's performance on a dataset from a different distribution than the training one is strongly correlated to the performance on the training set but with a constant offset.
\end{enumerate}
%
Since these scaling laws do not assume any special properties about the functions they are approximating, it is reasonable to expect these trends to carry over to SciML applications. While it is hard to find such studies done for physics-informed neural networks (architectures that can learn a given PDE solution function), a number of scaling studies are done in the field of operator learning, where many operators of interest are solution operators of PDEs. For example, the original DeepONet (deep operator network) paper~\citep{lu2019deeponet} showed various error convergence rates as a function of the amount of the network width, the amount of training data (number of functions), and the number of ``sensor locations'' (locations where the input functions are sampled).~\citep{de2022cost} reports power-law relationship between in- and out-of-distribution performance and the size of the training dataset for DeepONet, FNO, PCA-Net, and PARA-Net.~\citep{lanthaler2022error} studied the effect of the DeepONet network size on its approximation and generalization errors for specific operators. Finally,~\citep{liu2024neural} presents neural scaling laws (observed to be power laws) for DeepONets' approximation and generalization errors for general Lipschitz operators with respect to model and dataset sizes. While according to the scaling laws for operator learning methods the error can be driven down arbitrarily, the empirical results from~\citep{lu2019deeponet} show that this is not true. For example, Figure 6 in the paper shows that the relationship between training and test mean squared error with respect to the number of sensor locations initially follows a power law but plateaus after the number exceeds $10^1$. We see similar trends with the network width in Figure 2. This is consistent with the insights put forth by \citep{kaplan2020scaling}, that most scaling power laws plateau.
% \part One of the challenges as I understand in SciML applications is that your error requirements are much more stringent than those in ML applications. For certain kinds of ML models, there are so-called "scaling laws" (e.g., \url{https://arxiv.org/pdf/2001.08361}). Go over the paper above, summarizing the main results. Would you expect analogous results for SciML applications (e.g., learning very simple PDEs' solutions), and if so, do you expect to drive the error to an arbitrarily small quantity? This is an open ended question, so please include references to existing work.

{\bf Part b:} \citep{garg2022can} aims to study in-context learning of function classes with transformers. In-context learning is the ability of a model to, for a {\it previously unseen} function $f$, accurately predict $f(x_{i+1})$ at query point $x_{i+1}$ when given a prompt of $i$ ``in context'' sampling locations and samples of $f$, $\left(x_1, f(x_1), x_2, f(x_2), \dots, x_i, f(x_i) \right)$. For the class of linear functions, the transformer model can in-context learn well enough to be on par with several baselines; least squares estimator (known to be optimal), n-nearest neighbors, and averaging. In-context learning is also robust, it performs well even when the in-context outputs ($f(x_1), f(x_2), \dots, f(x_i)$) are noisy at inference time and when the distribution of the in-context examples and the query differ. The transformer model is also shown to in-context learn other function classes, mainly sparse linear functions, decision trees, and shallow neural networks. \citep{zhang2024trained} go a step further and show how in-context learning is ``learning a learning algorithm from data'' and study different distribution shifts under which the standard approach fails. Section 4.2 revisits three such shifts; task ($\D_{f}^{\text{train}} \neq \D_{f}^{\text{test}}$), query ($\D_{\text{query}}^{\text{test}} \neq \D_{x}^{\text{test}}$), and covariate ($\D_{x}^{\text{train}} \neq \D_{x}^{\text{test}}$). Transformers handle well the task and query shifts (the latter as long as $\D_{x}^{\text{train}} = \D_{x}^{\text{test}}$) but not covariate shifts. Interestingly, the same covariate shift phenomenon has been observed in SciML!

The original DeepONet paper~\citep{lu2019deeponet} states clearly that in order for the architecture to work, the sensor locations must be constant across all instances, essentially requiring $\D_{x}^{\text{train}} = \D_{x}^{\text{test}}$. This is generally true for most operator learning methods, though there is some work exploring ways to mitigate this by learning function encoders~\citep{ingebrand2025basis}. Most operator learning architectures also suffer from task shifts but none from query shifts (either at training or test time). When learning solution operators of PDEs, often the input functions are different initial and/or boundary conditions. As such, in order to avoid suffering from task shifts, the instances of input functions across training and test sets need to come from the same underlying distribution.
% \part Study one of the early papers on "in context" learning, specifically: \url{https://arxiv.org/pdf/2208.01066}. While SciML papers do not think about what they do as ICL, expecting a model to learn a solution to a ``new'' PDE is quite similar in spirit. Go over the paper above and summarize the main results. There have been many follow up works to the paper above claiming that the distributions are important, etc., but one interesting paper is: \url{https://arxiv.org/pdf/2306.09927}.
% \emph{Skim} the main results, but especially look at Section 4.2. Are you aware of analogous results in the SciML area? Does the distribution over boundary conditions, etc. matter for learnability?
\pagebreak
